---
title: "Aides Drug"
author: "Rachel Polack!"
date: "7/12/2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
females<-filter(triUmp, gender=="F")
```


```{r eval=FALSE}
females <- females %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

We need to separate out the females in the population, and run the interface against the subset that is only made up of females.
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

Given the two graphs above, we can see that the frequency of females who got aids is relatively much smaller compared to those who didn't. However, looking at the percentage graph we can see that among the females who got aids, most of these people have taken the drug. In other words, females who received the treatment got aids still. This may suggest the ineffectiveness of the drug against aids among females.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```
The two tables above show the specific values for the percentages of females who did and did not get the treatment and whether they got aids or not. Similar to the observations made in the graphical descriptive results, majority of the females, both who took the drug and did not, did get aids. Specifically, 96.31% of females who took the drug did not get aids while 97.58% of females who did not get the treatment didn't get aids. However, looking at the percentages of females who got aids, 58.82% took the drug which is higher than the percentage of females who received the placebo. Therefore, this supports the idea that the drug may not be effective against aids among females.




## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
Based on the chi-squared test above, the p-value is very small which means that we reject the null hypothesis which claims that there is no difference between taking the drug and taking the placebo for getting aids among females. This difference however, is opposite of the expected effect of the drug. Rather than prevent aids infection, the drug seems to cause it. Therefore, this drug does not appear effective against aids prevention.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In this write-up we analyzed whether the drug trialUmp is effective in preventing aids infection among females. This drug was compared to a placebo, and the frequency and percentage if infection were tallied. Based of the description and inferential results, it was observed among females who got aids, majority are those who got the treatment. In other words, the drug showed to be ineffective against the prevention of aids among females.


